Clustering Beat-Chroma Patterns in a Large Music Database

### 2010-03-23 Paper/Paper.tex
citations resolved
interpretation of fig 7

@@

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- i.e. the
harmony and melody.  As a computationally convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music typically remain recognizable
\cite{EllisPoliner-icassp07}.  To the extent, then, that beat-chroma
representations preserve tonal structure, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away some other aspects of the original audio, such as
instrumentation and other stylistic details.

@@

Prior work using small patches of chroma features includes ...
[casey \& slaney] [ellis \& poliner] [ellis ismir 07].

@@

\subsection{Codebook}
Visualizing codewords in a meaningful way is always a challenge, 
even if we have only a one or two hundreds of them.
We present $3$ figures in this section that should help the reader
foster his intuition. First, we display the most used codewords in
the testing set (Figure \ref{fig:codes1}). Then, we use the
 LLE\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}} 
algorithm \cite{Roweis2000} to visualize codewords arranged in
coherent groups (Figure \ref{fig:lle}).

>>>>

% Q: Why are there long spaces after "e.g." or "i.e."?
% A: LaTeX interprets a dot and a blank as the end of a sentence. To
%    prevent this, use "e.g.\ " and "i.e.\ " or "e.g.~" and "i.e.~".
\newcommand{\ie}{i.e.~}
\newcommand{\Ie}{I.e.~}
\newcommand{\eg}{e.g.~}
\newcommand{\Eg}{E.g.~}

@@

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- i.e. the
harmony and melody.  As a computationally convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music typically remain recognizable
\cite{Ellis2007a}.  To the extent, then, that beat-chroma
representations preserve tonal structure, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away some other aspects of the original audio, such as
instrumentation and other stylistic details.

@@

Prior work using small patches of chroma features includes ...
\cite{Casey2006,Ellis2007,Ellis2007a}.

@@

\subsection{Codebook}
Visualizing codewords in a meaningful way is always a challenge, even
if we have only a one or two hundreds of them.  We present $3$ figures
in this section that should help the reader foster his
intuition.

FIXME: fix the labels in the figure (label pitch class on the y axis,
and maybe only beats [0 2 4 6] on the ax axis, add codeword numbers to
title), etc.

The 25 most frequently used codewords in the testing set are shown in
Figure \ref{fig:codes1}.  The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 1),
perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
major and minor triads (code 6 and 5).  Many of the remaining
codewords correspond to transitions from one chord to another, \eg a
transition from a major triad to a minor triad in code 8.

(It is kind of interesting that sustained minor triad is more common
than the major triad (codes 4 and 5 in figure 7).  Don't fear the
reaper is in Amin, could it be related to this?)

Then, we use the LLE\footnote{implementation:
  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
algorithm \cite{Roweis2000} to visualize codewords arranged in
coherent groups (Figure \ref{fig:lle}).
